CMPSubset

### 2021-08-12 CMPSubset.Rmd
Update for index sort parameters

```{r}
seurat.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

@@

Add mitochondrial metadata and plot some basic features
```{r}
seurat.object[["percent.mt"]] <- PercentageFeatureSet(seurat.object, pattern = "^mt-")
VlnPlot(seurat.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r}
plot1 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(seurat.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```

@@

```{r}
print(paste("original object:", nrow(seurat.object@meta.data), "cells", sep = " "))
seurat.object <- subset(seurat.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000 & 
													percent.mt <= 5
												)
print(paste("new object:", nrow(seurat.object@meta.data), "cells", sep = " "))
```

```{r}
seurat.object <- NormalizeData(seurat.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features
```{r fig.width = 5, fig.height = 2}
seurat.object <- FindVariableFeatures(seurat.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(seurat.object), 10)
plot1 <- VariableFeaturePlot(seurat.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

@@

```{r echo = FALSE}
all.genes <- rownames(seurat.object)
seurat.object <- ScaleData(seurat.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
```

## Save raw object
```{r}
saveRDS(seurat.object, file = paste0(projectName, "_raw.RDS"))
```

```{r}
seurat.object <- RunPCA(seurat.object, features = VariableFeatures(seurat.object), ndims.print = 1:5, nfeatures.print = 5)
```

```{r}
DimPlot(seurat.object, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(seurat.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

@@

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(seurat.object, ndims = 50)
percent.variance(seurat.object@reductions$pca@stdev)
```
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

@@

Add new metadata column
```{r}
seurat.object@meta.data['clust.ID'] <- 0
head(seurat.object@meta.data)
```

now map new ids
```{r}
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust3.cells] <- 3
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust17.cells] <- 17
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust10.cells] <- 10
seurat.object@meta.data$clust.ID[rownames(seurat.object@meta.data) %in% clust11.cells] <- 11
```

do numbers make sense?
```{r}
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 10,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 11,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 17,])
nrow(seurat.object@meta.data[seurat.object@meta.data$clust.ID == 3,])
```

@@

# Total var 90%
## Neighborhood and umap
set total.var <- 90%
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 90))

seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
seurat.object <- FindClusters(seurat.object, resolution = 0.5)
seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)

saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```

```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(seurat.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

@@

Test: what percentage of each new clusterID matches one of the older clusters?
```{r}
for (meta.col in colnames(seurat.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		new.clusters <- sort(as.numeric(levels(seurat.object@meta.data[[meta.col]])))
		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
		colnames(enrich.df) <- c(3, 17, 10, 11)
		rownames(enrich.df) <- new.clusters
		meta.df <- seurat.object@meta.data
		for(row.id in rownames(enrich.df)){
			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
			for(col.id in colnames(enrich.df)){
				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
				pct.x <- as.integer(num.x / tot.clus *100)
				# print(pct.x)
				enrich.df[row.id, col.id] <- pct.x
			}
		}
		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
		print(enrich.df)
	}
}

@@

```{r fig.width = 4}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID", 
				# split.by = "orig.ident",
				cols = c("gray", "orange", "blue", "red", "green"),)
```
```{r fig.width = 4}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "orig.ident", 
				split.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"),)
```

# Total var 95%
## Neighborhood and umap
set total.var <- 95%
```{r}
tot.var <- percent.variance(seurat.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 95))

seurat.object <- FindNeighbors(seurat.object, dims = 1:ndims)
seurat.object <- FindClusters(seurat.object, resolution = 0.5)
seurat.object <- RunUMAP(seurat.object, dims = 1: ndims)

saveRDS(seurat.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	seurat.object <- FindClusters(seurat.object, resolution = x)
}
```

@@

```{r fig.width = 2}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "clust.ID", 
				pt.size = .1,
				# split.by = "orig.ident",
				cols = c("gray", "orange", "blue", "red", "green"),)
```
```{r fig.width = 4}
DimPlot(seurat.object,
				reduction = "umap",
				group.by = "orig.ident", 
				split.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"),)
```

### Gene expression of old clustrs on new map
Let's see if we can get some gene expression profiles on these...
```{r, fig.height=10, fig.width=18}
gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
VlnPlot(seurat.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
```

@@

calculate `FindAllMarkers()` for different idents and save to new file
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	Idents(seurat.object) <- tested.ident
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_res2.5.xlsx"), 
									 sheetName = tested.ident, 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
}
```

Create `FindAllMarkers()` lists for GSEA
```{r}
Idents(seurat.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(seurat.object)
```

@@

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	for (cluster in sort(as.numeric(levels(seurat.object@meta.data[[tested.ident]])))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
}
```

```{r}
for (cluster in sort(as.numeric(levels(seurat.object@meta.data$RNA_snn_res.2.5)))){
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_res2.5.xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
```

@@

# DGE w/ metadata against clust.ID against "0"
reset ident as "clust.ID" and rerun `FindAllMarkers()`
```{r}
	Idents(seurat.object) <- "clust.ID"
	all.markers <- FindAllMarkers(seurat.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_clustID.xlsx"), 
									 sheetName = "clustID", 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
```

```{r}
# Idents(seurat.object) <- "clust.ID"
for (cluster in unique(seurat.object@meta.data$clust.ID)){
	print(cluster)
	cluster.markers <- FindMarkers(seurat.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_clustID.xlsx"), 
									 sheetName = paste0("oldclust", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}

@@

# Distinguishing features of clusters
Previously defined biomark genes based on PC contributions. Original list was based on *all* msAggr, but let's see how CMP subset does?
```{r fig.height = 30, fig.width=6}
VizDimLoadings(seurat.object, dims = 1:10, nfeatures = 30, reduction = "pca", ncol = 2)
```

```{r}
pca.df <- seurat.object[["pca"]]
pca.df <- as.data.frame(as.matrix(slot(object = pca.df, name = "feature.loadings")))
print(seurat.object[["pca"]], dims = 2, nfeatures = 5)
rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2, decreasing = TRUE)[1:5], ])
rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2)[1:5], ])
```

@@

What about cell surface marker expression?
* Cd34
* Cd16/32
* Cd9
* Cd41
* Cd48
* Sca1 (just throw that in for sh*&s and giggles)
```{r, fig.height = 15, fig.width=10}
surface.markers <- c("Cd34", "Fcgr3", "Fcgr2b", "Cd9", "Itga2b", "Cd48", "Ly6a")
FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
```
Save as png
```{r}
png(filename = "FeaturePlot_CMP_surfaceMarkers_clustIDfacet.png", height = 1600, width = 1600)
FeaturePlot(seurat.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
dev.off()
```

@@

>>>>

```{r}
cmp.object<- CreateSeuratObject(counts = object.data, min.cells = 3, min.genes = 200, project = projectName)
```

@@

Add mitochondrial metadata and plot some basic features
```{r}
cmp.object[["percent.mt"]] <- PercentageFeatureSet(cmp.object, pattern = "^mt-")
VlnPlot(cmp.object, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, pt.size = 0, fill.by = 'orig.ident', )
```

```{r}
plot1 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = "orig.ident", pt.size = 0.01)
plot2 <- FeatureScatter(cmp.object, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = "orig.ident", pt.size = 0.01)
plot1 + plot2
```

@@

```{r}
print(paste("original object:", nrow(cmp.object@meta.data), "cells", sep = " "))
cmp.object <- subset(cmp.object, 
												subset = nFeature_RNA >=200 & 
													nFeature_RNA <= 4000 & 
													percent.mt <= 5
												)
print(paste("new object:", nrow(cmp.object@meta.data), "cells", sep = " "))
```

```{r}
cmp.object <- NormalizeData(cmp.object, normalization.method = "LogNormalize", scale.factor = 10000)
```

Find variable features
```{r fig.width = 5, fig.height = 2}
cmp.object <- FindVariableFeatures(cmp.object, selection.method = "vst", nfeatures = 2000)
top10 <- head(VariableFeatures(cmp.object), 10)
plot1 <- VariableFeaturePlot(cmp.object)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 + plot2

@@

```{r echo = FALSE}
all.genes <- rownames(cmp.object)
cmp.object <- ScaleData(cmp.object, features = all.genes, vars.to.regress = c("nFeature_RNA", "nCount_RNA"))
```

## Save raw object
```{r}
saveRDS(cmp.object, file = paste0(projectName, "_raw.RDS"))
```

```{r}
cmp.object <- RunPCA(cmp.object, features = VariableFeatures(cmp.object), ndims.print = 1:5, nfeatures.print = 5)
```

```{r}
DimPlot(cmp.object, reduction = "pca", group.by = "orig.ident")
VizDimLoadings(cmp.object, dims = 1:6, nfeatures = 10, reduction = "pca", ncol = 3)

@@

Calculate dimensionality
```{r, figures-side, fig.show='hold', out.width="50%"}
ElbowPlot(cmp.object, ndims = 50)
percent.variance(cmp.object@reductions$pca@stdev)
```
Number of PCs describing X% of variance

```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
paste0("Num pcs for 80% variance:", length(which(cumsum(tot.var) <= 80)))
paste0("Num pcs for 85% variance:", length(which(cumsum(tot.var) <= 85)))
paste0("Num pcs for 90% variance:", length(which(cumsum(tot.var) <= 90)))
paste0("Num pcs for 95% variance:", length(which(cumsum(tot.var) <= 95)))

@@

Add new metadata column
```{r}
cmp.object@meta.data['clust.ID'] <- 0
head(cmp.object@meta.data)
```

now map new ids
```{r}
cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust3.cells] <- 3
cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust17.cells] <- 17
cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust10.cells] <- 10
cmp.object@meta.data$clust.ID[rownames(cmp.object@meta.data) %in% clust11.cells] <- 11
```

do numbers make sense?
```{r}
nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 10,])
nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 11,])
nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 17,])
nrow(cmp.object@meta.data[cmp.object@meta.data$clust.ID == 3,])
```

@@

# Total var 90%
## Neighborhood and umap
set total.var <- 90%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 90))

cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
cmp.object <- FindClusters(cmp.object, resolution = 0.5)
cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)

saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
}
```

```{r}
for (meta.col in colnames(cmp.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(cmp.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```

for each resolution, number/percentage of cells in each cluster?

```{r}
tot.cells <- nrow(cmp.object@meta.data)
for (meta.col in colnames(cmp.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
		stats.df <- data.frame(matrix(ncol = 2, nrow = length(new.clusters)))
		colnames(stats.df) <- c("num_cells", "pct_pop")
		rownames(stats.df) <- new.clusters
		meta.df <- cmp.object@meta.data
		for(row.id in rownames(stats.df)){
				num.x <- nrow(meta.df[meta.df[meta.col] == row.id,])
				pct.x <- as.integer(num.x / tot.cells *100)
				# print(pct.x)
				stats.df[row.id, "num_cells"] <- num.x
				stats.df[row.id, "pct_pop"] <- pct.x
		}
		print(stats.df)
	}
}
```

@@

Test: what percentage of each new clusterID matches one of the older clusters?
```{r}
for (meta.col in colnames(cmp.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		new.clusters <- sort(as.numeric(levels(cmp.object@meta.data[[meta.col]])))
		enrich.df <- data.frame(matrix(ncol = 4, nrow = length(new.clusters)))
		colnames(enrich.df) <- c(3, 17, 10, 11)
		rownames(enrich.df) <- new.clusters
		meta.df <- cmp.object@meta.data
		for(row.id in rownames(enrich.df)){
			tot.clus <- nrow(meta.df[meta.df[[meta.col]] == row.id,])
			for(col.id in colnames(enrich.df)){
				num.x <- nrow(meta.df[(meta.df[[meta.col]] == row.id) & (meta.df$clust.ID == col.id),])
				pct.x <- as.integer(num.x / tot.clus *100)
				# print(pct.x)
				enrich.df[row.id, col.id] <- pct.x
			}
		}
		colnames(enrich.df) <- sapply(colnames(enrich.df), function(x) paste0("oldcluster", x))
		rownames(enrich.df) <- sapply(rownames(enrich.df), function(x) paste0("newcluster", x))
		xlsx::write.xlsx(enrich.df, file = paste0("PctOfNewClustersOverlappingOldClusters_", projectName, "_dim", ndims, ".xlsx"), sheetName = paste0(gsub("RNA_snn_", "", meta.col)), append = TRUE)
		print(enrich.df)
	}
}

@@

```{r fig.width = 4}
DimPlot(cmp.object,
				reduction = "umap",
				group.by = "clust.ID", 
				# split.by = "orig.ident",
				cols = c("gray", "orange", "blue", "red", "green"),)
```
```{r fig.width = 4}
DimPlot(cmp.object,
				reduction = "umap",
				group.by = "orig.ident", 
				split.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"),)
```

# Total var 95%
## Neighborhood and umap
set total.var <- 95%
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 95))

cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
cmp.object <- FindClusters(cmp.object, resolution = 0.5)
cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)

saveRDS(cmp.object, file = paste0(projectName, "_dim", ndims, ".RDS"))
```
Plot UMAP

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
}
```

@@

```{r fig.width = 2}
DimPlot(cmp.object,
				reduction = "umap",
				group.by = "clust.ID", 
				pt.size = .1,
				# split.by = "orig.ident",
				cols = c("gray", "orange", "blue", "red", "green"),)
```
```{r fig.width = 4}
DimPlot(cmp.object,
				reduction = "umap",
				group.by = "orig.ident", 
				split.by = "clust.ID",
				cols = c("gray", "orange", "blue", "red", "green"),)
```

### Gene expression of old clustrs on new map
Let's see if we can get some gene expression profiles on these...
```{r, fig.height=10, fig.width=18}
gene.list <- c("Gata1", "Gata2", "Pf4", "Dntt", "Mpo", "Meis1", "Irf8", "Elane", "Fli1", "Zfpm1")
VlnPlot(cmp.object, features = gene.list, group.by = "clust.ID", pt.size = 0.01, cols = c("gray", "orange", "blue", "red", "green"))
```

@@

calculate `FindAllMarkers()` for different idents and save to new file
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	Idents(cmp.object) <- tested.ident
	all.markers <- FindAllMarkers(cmp.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_res2.5.xlsx"), 
									 sheetName = tested.ident, 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
}
```

Create `FindAllMarkers()` lists for GSEA
```{r}
Idents(cmp.object) <- "RNA_snn_res.2.5"
res.2.5.allmarkers <- FindAllMarkers(cmp.object)
```

@@

calculate `FindMarkers()` that distinguish each cluster (might overlab between clusters)
```{r}
ident.list <- c("RNA_snn_res.0.5", "RNA_snn_res.1", "RNA_snn_res.1.5", "RNA_snn_res.2", "RNA_snn_res.2.5", "clust.ID")
for(tested.ident in ident.list){
	for (cluster in sort(as.numeric(levels(cmp.object@meta.data[[tested.ident]])))){
	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_", gsub("RNA_snn_", "", tested.ident), ".xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
}
```

```{r}
for (cluster in sort(as.numeric(levels(cmp.object@meta.data$RNA_snn_res.2.5)))){
	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_res2.5.xlsx"), 
									 sheetName = paste0("clst", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}
```

@@

# DGE w/ metadata against clust.ID against "0"
reset ident as "clust.ID" and rerun `FindAllMarkers()`
```{r}
	Idents(cmp.object) <- "clust.ID"
	all.markers <- FindAllMarkers(cmp.object)
	xlsx::write.xlsx(x = all.markers[,c("avg_log2FC", "p_val_adj", "cluster", "gene")], 
									 file = paste0(projectName, "_FindALLMarkers_clustID.xlsx"), 
									 sheetName = "clustID", 
									 col.names = TRUE, 
									 row.names = FALSE, 
									 append = TRUE)
```

```{r}
# Idents(cmp.object) <- "clust.ID"
for (cluster in unique(cmp.object@meta.data$clust.ID)){
	print(cluster)
	cluster.markers <- FindMarkers(cmp.object, ident.1 = cluster)
	xlsx::write.xlsx(x = cluster.markers[,c("avg_log2FC", "p_val_adj")], 
									 file = paste0(projectName, "_FindMarkers_clustID.xlsx"), 
									 sheetName = paste0("oldclust", cluster), 
									 col.names = TRUE, 
									 row.names = TRUE, 
									 append = TRUE)
}

@@

# Distinguishing features of clusters
Previously defined biomark genes based on PC contributions. Original list was based on *all* msAggr, but let's see how CMP subset does?
```{r fig.height = 30, fig.width=6}
VizDimLoadings(cmp.object, dims = 1:10, nfeatures = 30, reduction = "pca", ncol = 2)
```

```{r}
pca.df <- cmp.object[["pca"]]
pca.df <- as.data.frame(as.matrix(slot(object = pca.df, name = "feature.loadings")))
print(cmp.object[["pca"]], dims = 2, nfeatures = 5)
rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2, decreasing = TRUE)[1:5], ])
rownames(pca.df[pca.df$PC_2 %in% sort(pca.df$PC_2)[1:5], ])
```

@@

What about cell surface marker expression?
* Cd34
* Cd16/32
* Cd9
* Cd41
* Cd48
* Sca1 (just throw that in for sh*&s and giggles)
```{r, fig.height = 15, fig.width=10}
surface.markers <- c("Cd34", "Fcgr3", "Fcgr2b", "Cd9", "Itga2b", "Cd48", "Ly6a")
FeaturePlot(cmp.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
```
Save as png
```{r}
png(filename = "FeaturePlot_CMP_surfaceMarkers_clustIDfacet.png", height = 1600, width = 1600)
FeaturePlot(cmp.object, features = surface.markers, pt.size = 1, split.by = "clust.ID", ncol = 1)
dev.off()
```

@@

# Compare @ hierarchcial clusteirng

Do clustering using biomark RNAs as input
```{r}
# Read in BiomarkRNAs
biomark.rnas <- read.table('/Users/heustonef/Desktop/10XGenomicsData/BiomarkRNAs.txt')
biomark.rnas <- biomark.rnas$V1
```

use biomark RNAs to define dimensional reduction
```{r}
cmp.object <- readRDS("CMP_raw.RDS")
cmp.object <- RunPCA(cmp.object, features = biomark.rnas, ndims.print = 1:5, , nfeatures.print = 5)
ElbowPlot(cmp.object, ndims = 50)
```

Now run the clustering
```{r}
tot.var <- percent.variance(cmp.object@reductions$pca@stdev, plot.var = FALSE, return.val = TRUE)
ndims <- length(which(cumsum(tot.var) <= 90))

cmp.object <- FindNeighbors(cmp.object, dims = 1:ndims)
cmp.object <- FindClusters(cmp.object, resolution = 0.5)
cmp.object <- RunUMAP(cmp.object, dims = 1: ndims)

find the clusters

```{r}
for(x in c(0.5, 1, 1.5, 2, 2.5)){
	cmp.object <- FindClusters(cmp.object, resolution = x)
}
```

Plot the umaps and cell cluster ids
```{r}
for (meta.col in colnames(cmp.object@meta.data)){
	if(grepl(pattern = ("RNA_snn_res"), x = meta.col)==TRUE){
		myplot <- DimPlot(cmp.object, 
											group.by = meta.col,
											reduction = "umap", 
											cols = color.palette
											) + 
			ggtitle(paste0(projectName, " dim", ndims, "res", gsub("RNA_snn_res", "", meta.col) ))
		plot(myplot)
	}
}
```
### Calculate anticipated number of cells you'll find in each biomark cluster
Get # cells in each cluster

```{r}
tot.cellcount <- nrow(cmp.object@meta.data)
res05.list <- sort(unique(cmp.object@meta.data$RNA_snn_res.0.5), decreasing = FALSE)
sapply(res05.list, 
			 function(x){
			 	print(
			 		paste(
			 			"cluster", x, "=", 
			 			nrow(cmp.object@meta.data[cmp.object@meta.data$RNA_snn_res.0.5 == x,]), 
			 			"cells or", 
			 			round(nrow(cmp.object@meta.data[cmp.object@meta.data$RNA_snn_res.0.5 == x,])/tot.cellcount*100, digits = 2), 
			 			"% of total"
			 		)
			 	)
			 }
			)
```

So we did the dimensional reduction based on the biomark RNAs, then did our UMAP nearest neighbor clustering.

In the biomark hierarchcial clustering analysis I assayed 167 cells. The smallest cluster I detected had 3 cells, or 1.8% of total, and this is an uncomfortably small number of cells. Based on the UMAP calculations I would therefore expect to find 11 or 12 of the predicted 15 clusters. I found 12, and I don't really like that last one, so 11 or 12. Since I did the hierarchcial clustering yesterday and did this math today, we can say it was independent of these results and therefore totally legit. Yay!!
